### 2020-07-14 response/response-demography-3.tex

\begin{center} \textbf{\large Birth Spacing in the Presence of Son
Preference and 
Sex-Selective Abortions: India's Experience over Four Decades}
\end{center}

\begin{center} Response to Editor and Reviewer Comments \end{center}

\noindent Please find attached the revised version of my paper,
``Birth Spacing in the Presence of Son Preference and Sex-Selective
Abortions:
India's Experience over Four Decades.''

[Other changes]
\begin{enumerate}
	\item Rewritten Estimation Strategy section
	\begin{enumerate}
		\item Clarify proportionality assumption
		\item Better explain set-up of model and transformation of data needed
		\item Remove discussion of explanatory variables so they are
		only covered once (in the ``Data'' section)
		\item Better explanation of how I calculate measures used for the spacing
		analysis and all explanations gathered in this section
	\end{enumerate}
	\item Changes in Data section
	\begin{enumerate}
		\item Lead with what birth intervals I cover right after introducing
		data
		\item Move restriction to only Hindus to be the first restriction
		I discuss since that is the most important. 
		See also response to Review 2 below.
	\end{enumerate}
	\item Ensure consistent presentation of results
	\begin{enumerate}
		\item Use average rather than expected throughout now
		\item Update table notes to better describe calculation methods
	\end{enumerate}
\end{enumerate}

@@

Finally, I have made the discussion of the distribution of spacing a subsection
of the spacing section and combined the discussion of the results to another new 
subsection as well.
I hope this will make it easier for the reader to see the answer to the main
question of the paper.

TK I address the issue of the figure and table title above.

\item Please define sex selection early on.

I now define ``sex selection'' as ``the selective abortion of female fetuses 
based on prenatal sex determination'' in the second paragraph of the paper.

### 2020-07-21 paper/sex_selection_spacing-ver4.tex
Move to new combined graphs for spacing and mortality

\usepackage{amsmath}
\usepackage{fontspec}
\setromanfont[Ligatures=TeX]{TeX Gyre Pagella}
\usepackage{unicode-math}
\setmathfont{TeX Gyre Pagella Math}
\usepackage[title]{appendix}
\usepackage[margin=1.0in]{geometry}
\usepackage[figuresleft]{rotating}
\usepackage[longnamesfirst]{natbib}
\usepackage{dcolumn}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage[flushleft]{threeparttable}
\usepackage{setspace}
\usepackage[justification=centering]{caption}
\usepackage[font=scriptsize]{subfig}
\usepackage[xetex,colorlinks=true,linkcolor=black,citecolor=black,urlcolor=black]{hyperref}
\usepackage{adjustbox}
\usepackage{xfrac}
\usepackage{placeins}

@@

\date{June 2020}

@@

% Background
India experienced many positive changes over the last four decades:
the economy grew substantially,
schooling increased for both males and females,
and the total fertility rate fell to 2.2 
\citep{Bosworth2008,Dharmalingam2014,
International-Institute-for-Population-Sciences-IIPS2017}.

@@

% Questions addressed and why valuable
The main question here is: how did birth spacing respond to these changes, especially 
the spread of sex selection? 
The motivation is twofold.
First, prior research has failed to appreciate that each abortion increases the
interval between births by six to twelve months.%
\footnote{
The increase consists of three parts. 
First, after an abortion, the uterus needs at least two menstrual cycles to recover, 
or the likelihood of spontaneous abortion increases substantially \citep{zhou00b}. 
Second, the waiting time to conception is one to six months. 
Finally, sex determination tests are reliable only from three months of gestation. 
}
Hence, the growing use of sex-selective abortions may substantially increase birth spacing, 
but we know little about how much. 
Second, more female education, higher household income, and low and declining female labor 
force participation are all likely to impact birth spacing. 
The combined changes in birth spacing may, therefore, outpace what we have observed in 
other countries.

@@

The primary outcomes I examine are average birth intervals, the sex ratio of births, 
and the likelihood of a next birth. 
Because sex selection might also affect the distribution of birth intervals, I furthermore
examine the 25th and 75th percentiles of birth intervals. 
I estimate the model across four periods to capture the changing access and 
legality of sex selection. 
The key variables are maternal education, the sex composition of previous 
children, and area of residence.

@@

First, over the four decades, average birth intervals lengthen for all, but most for higher 
parities and more educated women. 
Women most likely to use sex selection---well-educated with no sons---had the most 
substantial lengthening of birth intervals and the most biased sex ratios. 
As a result, we now see cases that reverse the traditional spacing pattern, with some 
women with no sons having \emph{longer} birth intervals than those with sons.
Those least likely to use sex selection---less-educated in rural areas---still follow 
the traditional pattern of short spacing when they have girls and little evidence of sex 
selection.

@@

Figures \ref{fig:bootstrap_low} through \ref{fig:bootstrap_highest}
show the predicted outcomes for the four education levels by area of residence.%
\footnote{
For legibility, the graphs do not show standard errors.
The graphed values, with standard errors, are shown in Appendix Tables 
\ref{tab:avg_sex_ratio_low} through \ref{tab:avg_sex_ratio_highest}.
Furthermore, I do not show the results for the fourth spell the best-educated women since 
the low number of births, especially with one or more boys, make the spacing and sex ratio 
results very noisy.
Results are available upon request.
}
For comparison, the graphs for the sex ratio also show the natural sex ratio, 
approximately 51.2 percent \citep{ben-porath76b,jacobsen99,Portner2015b}.

% \begin{figure}[htpb]
% \centering
% \caption*{Urban}
% \subfloat[Second Spell]{\includegraphics[width=0.27\textwidth]{bs_spell2_low_urban_all}} 
% \subfloat[Third Spell]{\includegraphics[width=0.27\textwidth]{bs_spell3_low_urban_all}} 
% \subfloat[Fourth Spell]{\includegraphics[width=0.27\textwidth]{bs_spell4_low_urban_all}} 
% \\
% \caption*{Rural}
% \subfloat[Second Spell]{\includegraphics[width=0.27\textwidth]{bs_spell2_low_rural_all}} 
% \subfloat[Third Spell]{\includegraphics[width=0.27\textwidth]{bs_spell3_low_rural_all}} 
% \subfloat[Fourth Spell]{\includegraphics[width=0.27\textwidth]{bs_spell4_low_rural_all}} 
% \caption{low education}
% \label{fig:bootstrap_low}
% \end{figure}

% \captionsetup[subfigure]{position=top,captionskip=-1pt,farskip=-0.5pt}

% \begin{figure}[htpb]
% \centering
% \rotatebox[origin=c]{90}{\small{Urban}}
% \subfloat[Second Spell]{
%     \begin{minipage}{0.29\textwidth}
%         \includegraphics[width=\textwidth]{bs_spell2_low_urban_all}
%     \end{minipage}
% } 
% \subfloat[Third Spell]{
%     \begin{minipage}{0.29\textwidth}
%         \includegraphics[width=\textwidth]{bs_spell3_low_urban_all} 
%     \end{minipage}
% }
% \subfloat[Fourth Spell]{
%     \begin{minipage}{0.29\textwidth}
%         \includegraphics[width=\textwidth]{bs_spell4_low_urban_all}
%     \end{minipage}
% }
% \\
% \rotatebox[origin=c]{90}{\small{Rural}}
% \subfloat[Second Spell]{
%     \begin{minipage}{0.29\textwidth}
%         \includegraphics[width=\textwidth]{bs_spell2_low_rural_all}
%     \end{minipage}
% } 
% \subfloat[Third Spell]{
%     \begin{minipage}{0.29\textwidth}
%         \includegraphics[width=\textwidth]{bs_spell3_low_rural_all} 
%     \end{minipage}
% } 
% \subfloat[Fourth Spell]{
%     \begin{minipage}{0.29\textwidth}
%         \includegraphics[width=\textwidth]{bs_spell4_low_rural_all} 
%     \end{minipage}
% } 
% \caption{Estimated average spell length, sex ratio, and probability of 
% a next birth for women with no education}
% \label{fig:bootstrap_low}
% \end{figure}

\begin{sidewaysfigure}
\captionsetup[subfigure]{captionskip=-1pt,farskip=-0.5pt}
\centering
\rotatebox[origin=c]{90}{\small{Urban}}
\subfloat[Second Spell]{
    \begin{minipage}{0.31\textwidth}
        \includegraphics[width=\textwidth]{bs_spell2_low_urban_all}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.31\textwidth}
        \includegraphics[width=\textwidth]{bs_spell3_low_urban_all} 
    \end{minipage}
}
\subfloat[Fourth Spell]{
    \begin{minipage}{0.31\textwidth}
        \includegraphics[width=\textwidth]{bs_spell4_low_urban_all}
    \end{minipage}
}
\caption{Estimated average birth interval, sex ratio, and probability of 
a next birth for women with no education}
\label{fig:bootstrap_low}
\end{sidewaysfigure}

\begin{sidewaysfigure}
\centering
\ContinuedFloat
\rotatebox[origin=c]{90}{\small{Rural}}
\subfloat[Second Spell]{
    \begin{minipage}{0.31\textwidth}
        \includegraphics[width=\textwidth]{bs_spell2_low_rural_all}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.31\textwidth}
        \includegraphics[width=\textwidth]{bs_spell3_low_rural_all} 
    \end{minipage}
} 
\subfloat[Fourth Spell]{
    \begin{minipage}{0.31\textwidth}
        \includegraphics[width=\textwidth]{bs_spell4_low_rural_all} 
    \end{minipage}
} 
\caption{(Continued) Estimated average birth interval, sex ratio, and probability of 
a next birth for women with no education}
\label{fig:bootstrap_low}
\end{sidewaysfigure}

% \begin{figure}[htpb]
% \centering
% \caption*{Urban}
% \subfloat[Second Spell]{\includegraphics[width=0.27\textwidth]{bs_spell2_med_urban_all}} 
% \subfloat[Third Spell]{\includegraphics[width=0.27\textwidth]{bs_spell3_med_urban_all}} 
% \subfloat[Fourth Spell]{\includegraphics[width=0.27\textwidth]{bs_spell4_med_urban_all}} 
% \\
% \caption*{Rural}
% \subfloat[Second Spell]{\includegraphics[width=0.27\textwidth]{bs_spell2_med_rural_all}} 
% \subfloat[Third Spell]{\includegraphics[width=0.27\textwidth]{bs_spell3_med_rural_all}} 
% \subfloat[Fourth Spell]{\includegraphics[width=0.27\textwidth]{bs_spell4_med_rural_all}} 
% \caption{med education}
% \label{fig:bootstrap_med}
% \end{figure}

% \begin{figure}[htpb]
% \centering
% \rotatebox[origin=c]{90}{\small{Urban}}
% \subfloat[Second Spell]{
%     \begin{minipage}{0.29\textwidth}
%         \includegraphics[width=\textwidth]{bs_spell2_med_urban_all}
%     \end{minipage}
% } 
% \subfloat[Third Spell]{
%     \begin{minipage}{0.29\textwidth}
%         \includegraphics[width=\textwidth]{bs_spell3_med_urban_all} 
%     \end{minipage}
% }
% \subfloat[Fourth Spell]{
%     \begin{minipage}{0.29\textwidth}
%         \includegraphics[width=\textwidth]{bs_spell4_med_urban_all}
%     \end{minipage}
% }
% \\
% \rotatebox[origin=c]{90}{\small{Rural}}
% \subfloat[Second Spell]{
%     \begin{minipage}{0.29\textwidth}
%         \includegraphics[width=\textwidth]{bs_spell2_med_rural_all}
%     \end{minipage}
% } 
% \subfloat[Third Spell]{
%     \begin{minipage}{0.29\textwidth}
%         \includegraphics[width=\textwidth]{bs_spell3_med_rural_all} 
%     \end{minipage}
% } 
% \subfloat[Fourth Spell]{
%     \begin{minipage}{0.29\textwidth}
%         \includegraphics[width=\textwidth]{bs_spell4_med_rural_all} 
%     \end{minipage}
% } 
% \caption{Estimated average birth interval, sex ratio, and probability of 
% a next birth for women with one to seven years of education}
% \label{fig:bootstrap_med}
% \end{figure}

\begin{sidewaysfigure}
\centering
\rotatebox[origin=c]{90}{\small{Urban}}
\subfloat[Second Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{bs_spell2_med_urban_all}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{bs_spell3_med_urban_all} 
    \end{minipage}
}
\subfloat[Fourth Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{bs_spell4_med_urban_all}
    \end{minipage}
}
\caption{Estimated average birth interval, sex ratio, and probability of 
a next birth for women with one to seven years of education}
\label{fig:bootstrap_med}
\end{sidewaysfigure}

\begin{sidewaysfigure}
\centering
\ContinuedFloat
\rotatebox[origin=c]{90}{\small{Rural}}
\subfloat[Second Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{bs_spell2_med_rural_all}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{bs_spell3_med_rural_all} 
    \end{minipage}
} 
\subfloat[Fourth Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{bs_spell4_med_rural_all} 
    \end{minipage}
} 
\caption{(Continued) Estimated average birth interval, sex ratio, and probability of 
a next birth for women with one to seven years of education}
\label{fig:bootstrap_med}
\end{sidewaysfigure}

% \begin{figure}[htpb]
% \centering
% \caption*{Urban}
% \subfloat[Second Spell]{\includegraphics[width=0.27\textwidth]{bs_spell2_highest_urban_all}} 
% \subfloat[Third Spell]{\includegraphics[width=0.27\textwidth]{bs_spell3_highest_urban_all}} 
% \subfloat[Fourth Spell]{\includegraphics[width=0.27\textwidth]{bs_spell4_highest_urban_all}} 
% \\
% \caption*{Rural}
% \subfloat[Second Spell]{\includegraphics[width=0.27\textwidth]{bs_spell2_highest_rural_all}} 
% \subfloat[Third Spell]{\includegraphics[width=0.27\textwidth]{bs_spell3_highest_rural_all}} 
% \subfloat[Fourth Spell]{\includegraphics[width=0.27\textwidth]{bs_spell4_highest_rural_all}} 
% \caption{high education}
% \label{fig:bootstrap_high}
% \end{figure}

% \begin{figure}[htpb]
% \centering
% \rotatebox[origin=c]{90}{\small{Urban}}
% \subfloat[Second Spell]{
%     \begin{minipage}{0.29\textwidth}
%         \includegraphics[width=\textwidth]{bs_spell2_high_urban_all}
%     \end{minipage}
% } 
% \subfloat[Third Spell]{
%     \begin{minipage}{0.29\textwidth}
%         \includegraphics[width=\textwidth]{bs_spell3_high_urban_all} 
%     \end{minipage}
% }
% \subfloat[Fourth Spell]{
%     \begin{minipage}{0.29\textwidth}
%         \includegraphics[width=\textwidth]{bs_spell4_high_urban_all}
%     \end{minipage}
% }
% \\
% \rotatebox[origin=c]{90}{\small{Rural}}
% \subfloat[Second Spell]{
%     \begin{minipage}{0.29\textwidth}
%         \includegraphics[width=\textwidth]{bs_spell2_high_rural_all}
%     \end{minipage}
% } 
% \subfloat[Third Spell]{
%     \begin{minipage}{0.29\textwidth}
%         \includegraphics[width=\textwidth]{bs_spell3_high_rural_all} 
%     \end{minipage}
% } 
% \subfloat[Fourth Spell]{
%     \begin{minipage}{0.29\textwidth}
%         \includegraphics[width=\textwidth]{bs_spell4_high_rural_all} 
%     \end{minipage}
% } 
% \caption{Estimated average birth interval, sex ratio, and probability of 
% a next birth for women with eight to eleven years of education}
% \label{fig:bootstrap_high}
% \end{figure}

\begin{sidewaysfigure}
\centering
\rotatebox[origin=c]{90}{\small{Urban}}
\subfloat[Second Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{bs_spell2_high_urban_all}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{bs_spell3_high_urban_all} 
    \end{minipage}
}
\subfloat[Fourth Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{bs_spell4_high_urban_all}
    \end{minipage}
}
\caption{Estimated average birth interval, sex ratio, and probability of 
a next birth for women with eight to eleven years of education}
\label{fig:bootstrap_high}
\end{sidewaysfigure}

\begin{sidewaysfigure}
\centering
\ContinuedFloat
\rotatebox[origin=c]{90}{\small{Rural}}
\subfloat[Second Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{bs_spell2_high_rural_all}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{bs_spell3_high_rural_all} 
    \end{minipage}
} 
\subfloat[Fourth Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{bs_spell4_high_rural_all} 
    \end{minipage}
} 
\caption{(Continued) Estimated average birth interval, sex ratio, and probability of 
a next birth for women with eight to eleven years of education}
\label{fig:bootstrap_high}
\end{sidewaysfigure}

% \begin{figure}[htpb]
% \centering
% \caption*{Urban}
% \subfloat[Second Spell]{\includegraphics[width=0.27\textwidth]{bs_spell2_highest_urban_all}} 
% \subfloat[Third Spell]{\includegraphics[width=0.27\textwidth]{bs_spell3_highest_urban_all}} 
% \subfloat[Fourth Spell]{\includegraphics[width=0.27\textwidth]{bs_spell4_highest_urban_all}} 
% \\
% \caption*{Rural}
% \subfloat[Second Spell]{\includegraphics[width=0.27\textwidth]{bs_spell2_highest_rural_all}} 
% \subfloat[Third Spell]{\includegraphics[width=0.27\textwidth]{bs_spell3_highest_rural_all}} 
% \subfloat[Fourth Spell]{\includegraphics[width=0.27\textwidth]{bs_spell4_highest_rural_all}} 
% \caption{high education}
% \label{fig:bootstrap_high}
% \end{figure}

% \begin{figure}[htpb]
% \centering
% \rotatebox[origin=c]{90}{\small{Urban}}
% \subfloat[Second Spell]{
%     \begin{minipage}{0.29\textwidth}
%         \includegraphics[width=\textwidth]{bs_spell2_highest_urban_all}
%     \end{minipage}
% } 
% \subfloat[Third Spell]{
%     \begin{minipage}{0.29\textwidth}
%         \includegraphics[width=\textwidth]{bs_spell3_highest_urban_all} 
%     \end{minipage}
% }
% \\
% \rotatebox[origin=c]{90}{\small{Rural}}
% \subfloat[Second Spell]{
%     \begin{minipage}{0.29\textwidth}
%         \includegraphics[width=\textwidth]{bs_spell2_highest_rural_all}
%     \end{minipage}
% } 
% \subfloat[Third Spell]{
%     \begin{minipage}{0.29\textwidth}
%         \includegraphics[width=\textwidth]{bs_spell3_highest_rural_all} 
%     \end{minipage}
% } 
% \caption{Estimated average birth interval, sex ratio, and probability of 
% a next birth for women with twelve or more years of education}
% \label{fig:bootstrap_highest}
% \end{figure}

\begin{sidewaysfigure}
\centering
\rotatebox[origin=c]{90}{\small{Urban}}
\subfloat[Second Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{bs_spell2_highest_urban_all}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{bs_spell3_highest_urban_all} 
    \end{minipage}
}
\caption{Estimated average birth interval, sex ratio, and probability of 
a next birth for women with twelve or more years of education}
\label{fig:bootstrap_highest}
\end{sidewaysfigure}

\begin{sidewaysfigure}
\centering
\ContinuedFloat
\rotatebox[origin=c]{90}{\small{Rural}}
\subfloat[Second Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{bs_spell2_highest_rural_all}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{bs_spell3_highest_rural_all} 
    \end{minipage}
} 
\caption{(Continued) Estimated average birth interval, sex ratio, and probability of 
a next birth for women with twelve or more years of education}
\label{fig:bootstrap_highest}
\end{sidewaysfigure}

@@

In the latest period, the shortest average birth interval is 33 months for
women with no education in the second spell, an increase of a couple of
months from the first period.
The longest average interval was 51 months for the third spell for urban women with twelve or 
more years of education who had no boys, an increase of approximately a year.
For comparison, the median birth intervals reported for the NFHS have barely
moved over time, staying at approximately 31 months.%
\footnote{
See
\citet[][p.\ 110--112]{International-Institute-for-Population-Sciences-IIPS1995},
\citet[][p.\ 98--103]{International-Institute-for-Population-Sciences-IIPS2000},
\citet[][p.\ 88--91]{International-Institute-for-Population-Sciences-IIPS2007}, and
\citet[][p.\ 81--82]{International-Institute-for-Population-Sciences-IIPS2017}.
Median estimates are in
Appendix Tables \ref{tab:p25_p50_p75_low} through \ref{tab:p25_p50_p75_highest}.
}
This lack of change in the NFHS reported birth intervals underscores the importance of 
accounting for censoring of birth spells when trying to understand how birth spacing 
is changing over time.

@@

\begin{figure}[htpb]
\captionsetup[subfigure]{labelformat=empty,position=top}
\captionsetup{font=footnotesize,skip=2pt}
\centering
\caption*{No education}
\subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{combined_p25_p75_spell2_low_urban}}
\subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{combined_p25_p75_spell3_low_urban}}
\subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{combined_p25_p75_spell4_low_urban}}
\\
\caption*{One to Seven Years of Education}
\subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{combined_p25_p75_spell2_med_urban}}
\subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{combined_p25_p75_spell3_med_urban}}
\subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{combined_p25_p75_spell4_med_urban}}
\\
\caption*{Eight to Eleven Years of Education}
\subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{combined_p25_p75_spell2_high_urban}}
\subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{combined_p25_p75_spell3_high_urban}}
\subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{combined_p25_p75_spell4_high_urban}}
\\
\caption*{Twelve or More Years of Education}
\subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{combined_p25_p75_spell2_highest_urban}}
\subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{combined_p25_p75_spell3_highest_urban}}
\begin{minipage}{0.32\textwidth}\hspace{1cm}\end{minipage}
\captionsetup{font=normalsize}
\caption{
25th and 75th Percentile Birth Intervals for Urban Women by Spell and Education
}
\label{fig:combined_p25_p75_urban}
\end{figure}

\begin{figure}[htpb]
% \captionsetup[subfigure]{labelformat=empty,position=top}
\captionsetup[subfigure]{position=top}
\captionsetup{font=footnotesize,skip=2pt}
\centering
\caption*{No education}
\subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{combined_p25_p75_spell2_low_rural}}
\subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{combined_p25_p75_spell3_low_rural}}
\subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{combined_p25_p75_spell4_low_rural}}
\\
\caption*{One to Seven Years of Education}
\subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{combined_p25_p75_spell2_med_rural}}
\subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{combined_p25_p75_spell3_med_rural}}
\subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{combined_p25_p75_spell4_med_rural}}
\\
\caption*{Eight to Eleven Years of Education}
\subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{combined_p25_p75_spell2_high_rural}}
\subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{combined_p25_p75_spell3_high_rural}}
\subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{combined_p25_p75_spell4_high_rural}}
\\
\caption*{Twelve or More Years of Education}
\subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{combined_p25_p75_spell2_highest_rural}}
\subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{combined_p25_p75_spell3_highest_rural}}
\begin{minipage}{0.32\textwidth}\hspace{1cm}\end{minipage}
\captionsetup{font=normalsize}
\caption{
25th and 75th Percentile Birth Intervals for Rural Women by Spell and Education
}
\label{fig:combined_p25_p75_rural}
\end{figure}

@@

The final group consists of those women, who, in the absence of a son, have the most 
biased sex ratios and the longest birth intervals, and where this spacing is longer than 
for the other possible sex compositions. 
The most extreme example is best-educated urban women with two girls, who have a sex ratio 
approaching 80 and an average birth interval of 42 months. 
This birth interval is seven months longer than with at least one boy.

>>>>

\usepackage{amsmath}
\usepackage{fontspec}
\setromanfont[Ligatures=TeX]{TeX Gyre Pagella}
\usepackage{unicode-math}
\setmathfont{TeX Gyre Pagella Math}
\usepackage[title]{appendix}
\usepackage[margin=1.0in]{geometry}
\usepackage[figuresleft]{rotating}
\usepackage[longnamesfirst]{natbib}
\usepackage{dcolumn}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage[flushleft]{threeparttable}
\usepackage{setspace}
\usepackage[justification=centering,skip=2pt]{caption}
\usepackage[font=scriptsize]{subfig}
\usepackage[xetex,colorlinks=true,linkcolor=black,citecolor=black,urlcolor=black]{hyperref}
\usepackage{adjustbox}
\usepackage{xfrac}
\usepackage{placeins}

@@

\date{July 2020}

@@

% Background
India experienced many positive changes over the last four decades:
the economy grew substantially,
schooling increased for males and females,
and the total fertility rate fell to 2.2 
\citep{Bosworth2008,Dharmalingam2014,
International-Institute-for-Population-Sciences-IIPS2017}.

@@

% Questions addressed and why valuable
The main question here is: how did birth spacing respond to these changes, especially 
the spread of sex selection? 
The motivation is twofold.
First, prior research has failed to appreciate that each abortion increases the
interval between births by six to twelve months.%
\footnote{
The increase consists of three parts. 
First, after an abortion, the uterus needs at least two menstrual cycles to recover, 
or the likelihood of spontaneous abortion increases substantially \citep{zhou00b}. 
Second, the waiting time to conception is one to six months. 
Finally, sex determination tests are reliable only from three months of gestation. 
}
Hence, the growing use of sex-selective abortions may substantially increase birth spacing, 
but we know little about how much. 
Second, more female education, higher household income, and low and declining female labor 
force participation all likely impact birth spacing. 
The combined changes in birth spacing may, therefore, outpace what we have observed in 
other countries.

@@

The primary outcomes I examine are average birth intervals, the sex ratio of births, 
and the likelihood of giving birth. 
Because sex selection might also affect the distribution of birth intervals, I furthermore
examine the 25th and 75th percentiles of birth intervals. 
I estimate the model across four periods to capture the changing access and 
legality of sex selection. 
The key variables are maternal education, the sex composition of previous 
children, and area of residence.

@@

First, average birth intervals lengthen over the four decades, more so the higher the 
parity and the more educated the woman.
Women most likely to use sex selection---well-educated with no sons---had the most 
substantial lengthening of birth intervals and the most biased sex ratios. 
As a result, we now see cases that reverse the traditional spacing pattern, with some 
women with no sons having \emph{longer} birth intervals than those with sons.
Those least likely to use sex selection---less-educated in rural areas---still follow 
the traditional pattern of short spacing when they have girls and little evidence of sex 
selection.

@@

Figures \ref{fig:spacing_low_rural} through \ref{fig:spacing_highest_urban}
show the predicted outcomes for the four education levels by area of residence.%
\footnote{
For legibility, the graphs do not show standard errors.
The graphed values, with standard errors, are shown in Appendix Tables 
\ref{tab:avg_sex_ratio_low} through \ref{tab:avg_sex_ratio_highest}.
Furthermore, I do not show the results for the fourth spell the best-educated women since 
the low number of births, especially with one or more boys, make the spacing and sex ratio 
results very noisy.
Results are available upon request.

[Urban low and rural highest in online Appendix]

}
For comparison, the graphs for the sex ratio also show the natural sex ratio, 
approximately 51.2 percent \citep{ben-porath76b,jacobsen99,Portner2015b}.

% \begin{figure}
% \centering
% \includegraphics[width=\textwidth,height=\textheight,keepaspectration=true]{bs_low_urban}
% \caption{combined graphs - low urban}
% \label{fig:spacing_low_urban}
% \end{figure}

\begin{figure}
\centering
\includegraphics[width=\textwidth,height=\textheight,keepaspectration=true]{bs_low_rural}
\caption{combined graphs - low rural}
\label{fig:spacing_low_rural}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=\textwidth,height=\textheight,keepaspectration=true]{bs_med_urban}
\caption{combined graphs - med urban}
\label{fig:spacing_med_urban}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=\textwidth,height=\textheight,keepaspectration=true]{bs_med_rural}
\caption{combined graphs - med rural}
\label{fig:spacing_med_rural}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=\textwidth,height=\textheight,keepaspectration=true]{bs_high_urban}
\caption{combined graphs - high urban}
\label{fig:spacing_high_urban}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=\textwidth,height=\textheight,keepaspectration=true]{bs_high_rural}
\caption{combined graphs - high rural}
\label{fig:spacing_high_rural}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=\textwidth,height=\textheight,keepaspectration=true]{bs_highest_urban}
\caption{combined graphs - highest urban}
\label{fig:spacing_highest_urban}
\end{figure}

% \begin{figure}
% \centering
% \includegraphics[width=\textwidth,height=\textheight,keepaspectration=true]{bs_highest_rural}
% \caption{combined graphs - highest rural}
% \label{fig:spacing_highest_rural}
% \end{figure}

@@

In the latest period, the shortest average birth interval is 33 months for
women with no education in the second spell, an increase of a couple of
months from the first period.
The longest average interval was 51 months for the third spell for urban women with twelve 
or more years of education who had no boys, an increase of approximately a year over
4 decades.
For comparison, the median birth intervals reported for the NFHS have barely
moved over time, staying at approximately 31 months, and fast changing countries have 
median birth intervals that increase change around 3.5 months per decade 
\citep{Rutstein2011,Casterline2016}.%
\footnote{
Median estimates are in Appendix Tables \ref{tab:p25_p50_p75_low} through 
\ref{tab:p25_p50_p75_highest}.
}
This lack of change in the NFHS reported birth intervals over round underscores the 
importance of accounting for censoring of birth spells when trying to understand how birth 
spacing is changing over time.

@@

@@

The final group consists of those women, who, in the absence of a son, have the most 
biased sex ratios and the longest birth intervals, and where this spacing is longer than 
for the other possible sex compositions. 
The most extreme example is best-educated urban women with two girls, who have a sex ratio 
approaching 80 and an average birth interval of 51 months. 
This birth interval is seven months longer than with at least one boy.

Even more striking is the change for this group when sex selection was first introduced.
There was a six months jump, from 39.5 to 45.4, in average spacing between the first and 
the second period for women with two girls, but only a month increase for the other sex 
compositions.
Furthermore, the 75th percentile birth interval with two girls increased from 48 months to 
64 months with two girls over the period, while the other sex compositions showed a 
slightly decrease from around 55 months to 54 months.

Theses changes may even be an underestimate of the changes in birth spacing sine
this particular group appear to have had access to sex selection even before it became
widespread as shown by the unequal sex ratio for the 1972--1984 period for women with
two girls.
